---
title: "Leiden Example"
author: "Brandt Bessell"
date: "2023-12-08"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

To run our package, first install it and some other dependencies we will use to
benchmark

```{r}
if (!require("AluNet")) {
  install.packages("AluNet")
}
if (!require("microbenchmark")) {
  install.packages("microbenchmark")
}
```

Then we will load it, igraph, and microbenchmark to benchmark against the 
current gold standard of clustering algorithm implementations.

```{r}
library(Rcpp)
library(igraph)
library(microbenchmark)
#library("AluNet")
```

We provide some enviromental context for our tests in case you have trouble 
compiling the cpp files.
```{r}
setwd("C:/Users/bbessell/Projects/GitHub/AluNet")
Sys.setenv(PKG_CXXFLAGS = "-std=c++17")
Sys.which("make")
Sys.which("g++")
sourceCpp("rcpp/Leiden.cpp", verbose = F, rebuild = TRUE)
```

First, we check against several toy models. Note that the RNG is different for
each implementation of Leiden so we will run several different seeds for each
one to show that they are functioning similarly.

To do this we define a function to to run igraph's Leiden algorithm and plot 
the output, we define it here since it is not part of our package. 

```{r}
#gdf <- make_graph("noperfectmatching")

runIgraph <- function(gdf, iterations, gamma, theta) {

  ldc <- cluster_leiden(
    gdf,
    resolution_parameter=gamma,
    objective_function="CPM",
    weights=g_list[["weight"]],
    beta=theta,
    n_iterations=iterations,
    vertex_weights = rep(1, length(unique(c(g_list[["to"]], g_list[["from"]]))))
  )
  
  return(ldc)
}

plotIgraph <- function(gdf, ldc) {
  color_palette <- heat.colors(100)
  
    # create colormap for the communities
    get_community_colors <- function(num_communities) {
    rainbow(num_communities)
    }
  
    # map communities to the colormap
    num_communities <- length(unique(ldc$membership))
    community_colors <- get_community_colors(num_communities)
    colors <- community_colors[as.factor(ldc$membership)]
  
  plot(
    gdf,
    layout = layout_with_fr(gdf),
    vertex.color = colors,  # Color nodes based on community
    #edge.color = color_palette[g_list$weight],  # Map edge color from cold to hot
    vertex.label = V(gdf)$community,
    edge.arrow.size = 0,  # Set arrow size to 0 to remove arrows
    main = "iGraph's C-based clustering"
  )
  
  # Add a legend for community colors
  legend("topright", legend=unique(ldc$membership),
         col = community_colors, pch = 19,
         title = "Communities", cex = 0.8)
}
```

Our first toy model is "noperfectmatching" which has very well defined
communities except for the central node which can belong to any of them
depending on the starting conditions. This is a simple model and thus should 
converge in only a single iteration, despite us setting a maximum of 10.

We define our resolution parameter, gamma as .1 which should be small enough
to identify more granular communities that will make it easier to compare its
accuracy to igraph, we set theta to 0.01 because it is the middle of the 
recommended ranges of theta by the authors of the Leiden algorithm which should
result in a moderate degree of randomness in the refine step. Since the toy
models come from the igraph package, we also have to convert them into a format
interpretable by our function, which is a list of connections "to" "from" and
"weight."

```{r}
toy = "noperfectmatching"
iterations <- 10
gamma <- .1  # gamma > 0
theta <- 0.01 # good theta is 0.005 < 0.05

g <- as_edgelist(make_graph(toy))
g_list <- list(
  to=g[,1],from=g[,2],
  weight=rep(1, length(g[,1]))  # all 1
  
)
```

```{r}
seed <- 1
result <- runLeiden(g_list, iterations, gamma, theta, seed)
plotLeiden(result)
```


```{r}
gdf <- make_graph(toy)
set.seed(1)
igraph_result <- runIgraph(gdf, iterations, gamma, theta)
plotIgraph(gdf, igraph_result)
```

Next we benchmark the model to show that igraph, as expected, outperforms our
implementation due to its lower level control over memory, and use of vectors
instead of hash maps. 

```{r}
rm(.Random.seed)
# Use microbenchmark to time the function
mb_ours <- microbenchmark(
  runLeiden(g_list, 1, gamma, theta),
  times=10
)

mb_igraph <- microbenchmark(
  runIgraph(gdf, 1, gamma, theta),
  times=10
)
# Print the result
print(rbind(mb_ours, mb_igraph))
```

Next we try with some harder toy models, 

```{r}
toy = "Zachary"
iterations <- 10
gamma <- .1  # gamma > 0
theta <- 0.01 # good theta is 0.005 < 0.05

g <- as_edgelist(make_graph(toy))
g_list <- list(
  to=g[,1],from=g[,2],
  weight=rep(1, length(g[,1]))  # all 1
  
)
gdf <- make_graph(toy)
```

```{r}
seed <- 1
result <- runLeiden(g_list, iterations, gamma, theta, seed)
plotLeiden(result)
```



```{r}


rm(.Random.seed)
# Use microbenchmark to time the function
mb_ours <- microbenchmark(
  runLeiden(g_list, 1, gamma, theta),
  times=10
)

mb_igraph <- microbenchmark(
  runIgraph(gdf, 1, gamma, theta),
  times=10
)
# Print the result
print(rbind(mb_ours, mb_igraph))
```

To show its utility on biological data, we first download some Fit-Hi-C data (cite)

```{r, eval=FALSE}
library(R.utils)
file_path <- "path/to/your/working/directory"
setwd(file_path)
file_name <- "hic_data.txt.gz"
url <- "https://www.ncbi.nlm.nih.gov/geo/download/?acc=GSE148434&format=file&file=GSE148434%5FNS%5Fall%2E5kb%2E1e%2D2%2EFitHiC%2Eall%2Etxt%2Egz"

download.file(url, paste(file_path, file_name, sep = ""), mode = "wb")
gunzip(file_name, remove=FALSE)
```

Then we download Dfam mobile elements (cite) (be careful, it is very large)

```{r, eval=FALSE}


```

Then we download some example ATAC-seq data (cite)

```{r, eval=FALSE}

```
